Galaxy Zoo: Morphological Classifications for 120,000 Galaxies in HST Legacy Imaging

### 2016-05-31 writeup/gz_hubble_data.tex
7.3 is now organized by survey

@@

@@

@@

The AEGIS sample had only 485~galaxies matched with GZH, and so the statistical
uncertainties in the analysis are higher. However, at \fbest$>0.8$ there is a
clear separation of featured and unfeatured galaxies, though this is not seen
at lower \fbest.  GOODS galaxies, which are the only comparison sample that
also uses visual morphology, shows a very similar trend, with nearly 100\% pure
samples at the lowest and highest values of \fbest.

Galaxies from GEMS (which
are only split into two types) also have a strong correlation with \fbest,
which is somewhat surprising (but encouraging) considering the known limits of
assigning galaxy morphology based only on a single \sersic{} parameter. 
Visual inspection of images of the exceptions --- where both \fbest{} and $n$
are high --- show that most are obvious spirals but with prominent bulges. This
means that the single-component \sersic{} fit is likely choosing too small of
an effective radius and missing the extended disk structure for a large
population of galaxies. This is a known issue with using $n$ as a standalone
measure of galaxy morphology -- it is relatively common, for example, to
measure high $n$ for an object with an exponential disk but a central cusp in
the light profile (eg, due to an AGN, cluster, foreground star, noise, etc).

Galaxies from the three COSMOS morphology catalogues all have morphologies that are
reasonably well-predicted by the GZH based on \fbest. The relation is strongest
for early-type galaxies, since the comparable fractions of irregulars and
spirals shows that clumpy/asymmetric morphologies are clearly also being
captured by the \fbest{} parameter. The small dropoff in the fraction of
early-types for the Cassata and ZEST samples at \fbest$<0.1$ is somewhat
puzzling, but could potentially be due to the presence of a small
population of extremely compact objects (stars or quasars) that would be
classified as artifacts in GZH, explaining the dip in \fbest.

%We also emphasise that the previously published COSMOS catalogues are not
%internally consistent; for example, only 64\% of the galaxies in \citet{tas11}
%have the same morphological class as \citet{cas07}. The bulk of these are
%galaxies classified as ``spiral'' using one technique and ``irregular'' using
%another (\citealt{cas07} is more likely to classify galaxies as irregular).  A
%detailed comparison of the external catalogues is outside the scope of this
%paper, but we emphasise only that it is \emph{impossible} for any catalogue to be
%consistent with all three.  We interpret the broad agreements between
%early-types, spirals, and irregulars for all three samples as a validation of
%the crowdsourced visual classifications provided by GZH.

%\begin{figure*}
%\center
%\includegraphics[width=1.0\textwidth]{figures/comparisons_odd.pdf}
%\caption{Same as Figure~\ref{fig:comparisons_features}, but comparing external
%morphologies to \fodd{} in GZH. Measurement of \fodd{} excludes galaxies
%labeled as ``odd'' due to lenses or dust lanes ($<1\%$ of the total), which are GZH
%categories not expected to correlate with larger-scale structural parameters.}
%\label{fig:comparisons_odd}
%\end{figure*}
%
%Figure~\ref{fig:comparisons_odd} analyses the same sets of matched galaxies, but
%compares the data to the \fodd{} parameter in GZH to examine its correlation with
%irregular morphologies. The relationship in this case is less clear; for the
%COSMOS catalogues, in particular, early-types, spirals, and irregulars are
%roughly equally probable at \fodd$>0.5$. In both AEGIS and GOODS (there is no
%measurement of irregular structure for GEMS in \citealt{hau07}), the proportion
%of irregular galaxies does increase with \fodd. However, the actual number of
%galaxies at \fodd$>0.5$ is quite small ($N\lesssim100$) in both samples, and so
%estimating an accurate fraction is less precise.  We also note that this
%particular question ({\it ``anything odd?''}) has somewhat uncertain phrasing
%in GZH, and that the design of the tree explicitly separates clumpy from
%irregular or merging morphologies, which would have similar properties in
%automatic techniques. A more accurate selection of irregular morphologies in
%GZH should use a weighted combination of \fbest, \fclumpy, and \fodd.

A detailed comparison of the external catalogues is outside the scope of this
paper, but we emphasise only that it is impossible for any catalogue to be fully
consistent with all of the previously published datasets (for example, only
64\% of the galaxies in \citealt{tas11} have the same morphological class as
the same galaxies in \citealt{cas07}). We interpret the broad agreements
between early-types, spirals, and irregulars between GZH and all three methods
of automated classification as a validation of both the crowdsourced visual
classifications provided by GZH and those automated methods.

>>>>

The AEGIS sample had only 485~total galaxies matched with GZH, and so the
statistical uncertainties in the analysis are higher. However, at \fbest$>0.8$
there is a clear separation of featured and unfeatured galaxies in
Figure~\ref{fig:comparisons_features}, though this is not seen at lower \fbest.

@@

Galaxies from GEMS also have a strong correlation with \fbest{}
(Figure~\ref{fig:comparisons_features}), which is somewhat surprising (but
encouraging) considering the known limits of assigning galaxy morphology based
only on a single \sersic{} parameter.  Visual inspection of images of the
exceptions --- where both \fbest{} and $n$ are high --- show that most are
obvious spirals but with prominent bulges. This means that the single-component
\sersic{} fit is likely choosing too small of an effective radius and missing
the extended disk structure for a large population of galaxies. This is a known
issue with using $n$ as a standalone measure of galaxy morphology -- it is
relatively common, for example, to measure high $n$ for an object with an
exponential disk but a central cusp in the light profile (eg, due to an AGN,
cluster, foreground star, noise, etc).

@@

The GOODS galaxies show a very similar trend to the AEGIS images in
Figure~\ref{fig:comparisons_features}, with nearly 100\% pure samples at the
lowest and highest values of \fbest. We conjecture that the strong agreement
is at least partially due to the shared method of visual classification,
and reinforces the findings from several phases of Galaxy~Zoo that crowdsourced
classifications can be competitive with dedicated experts.

@@

Galaxies from the COSMOS morphology catalogues all have morphologies that are
reasonably well-predicted by the GZH based on \fbest. The relation is strongest
for early-type galaxies, since the comparable fractions of irregulars and
spirals shows that clumpy/asymmetric morphologies are clearly also being
captured by the \fbest{} parameter. The small dropoff in the fraction of
early-types for the Cassata and ZEST samples at \fbest$<0.1$ in
Figure~\ref{fig:comparisons_features} is somewhat puzzling, but could
potentially be due to the presence of a small population of extremely
compact objects (stars or quasars) that would be classified as artifacts in
GZH, explaining the dip in \fbest.

A detailed comparison of all external catalogues is outside the scope of this
paper, but we emphasise only that it is impossible for any catalogue to be fully
consistent with all of the previously published datasets (for example, only
64\% of the galaxies in \citealt{tas11} have the same morphological class as
the same galaxies in \citealt{cas07}). We interpret the broad agreements
between early-types, spirals, and irregulars between GZH and all three methods
of automated classification as a validation of both the crowdsourced visual
classifications provided by GZH and those automated methods.
